Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone

### 2020-06-30 paper/ifo-appendix.tex
winter comments part 2

The intensity $I(t)$ at $P$ is then the norm squared of the total electric field amplitude,
\begin{equation}
I(t) = \lvert A e^{i (k L - \omega t)}  + A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \rvert^2\,.
\label{eqn:intensity1}    
\end{equation}
Some elementary trigonometry yields
\begin{equation}
%I(t) = 2 A^2 (1 + \cos{\phi})
I(t) = 2 A^2 \left[ 1 + \cos{\left( 2 k d(t) \cos{\theta} \right) } \right]  \label{eqn:intensity2}\,.
\end{equation}
Equation~\ref{eqn:intensity2} shows that $I(t)$ and $d(t)$ are related nonlinearly. 
Near a bright fringe, where one has $\cos{[2kd\cos{(\theta)}]} \approx 1$, the nonlinearity is quadratic, with $\delta I \propto (\delta d)^2$ for small $\delta d$. 
Consequently, an oscillation of the form $\delta d \propto \exp{( -i \Omega t )}$ in the arm length difference produces a DC signal and a frequency-double signal $\propto \exp{(-2i\Omega t)}$ in $\delta I$. 
This is discussed further in Section~\ref{}\han{to do}.

@@

@@

\han{to do! psudo code here}
%We consider a spectrogram grid where each element in the grid is the Fourier amplitude of a particular frequency at a particular time. These are normalised between $(0, 1)$ to use as multiplicative weights. In fact, we take the logarithm and so the sum of all weights to avoid underflow. The best path has the highest weight among all paths.

@@

Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram. 
It is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}. A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}. This design was based off standard examples of photo-detectors, connecting an ADC to a Pi, and Sallen-Key filters.

>>>>

The intensity $I(t)$ at $P$ is then the norm squared of the total electric field amplitude,
\begin{equation}
I(t) = \lvert A e^{i (k L - \omega t)}  + A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \rvert^2\,.
\label{eqn:intensity1}    
\end{equation}
Some elementary trigonometry yields
\begin{equation}
%I(t) = 2 A^2 (1 + \cos{\phi})
I(t) = 2 A^2 \left[ 1 + \cos{\left( 2 k d(t) \cos{\theta} \right) } \right]  \label{eqn:intensity2}\,.
\end{equation}
Equation~\ref{eqn:intensity2} shows that $I(t)$ and $d(t)$ are related nonlinearly. 
Near a bright fringe, where one has $\cos{[2kd\cos{(\theta)}]} \approx 1$, the nonlinearity is quadratic, with $\delta I \propto (\delta d)^2$ for small $\delta d$. 
Consequently, an oscillation of the form $\delta d \propto \exp{( -i \Omega t )}$ in the arm length difference produces a DC signal and a frequency-double signal $\propto \exp{(-2i\Omega t)}$ in $\delta I$. 
This is discussed further in Section~\ref{}\han{to do}. \jam{[not quite sure how much further we can take discussion about this, are we wanting results that show the doubling?]}

@@

\section{Likelihood of sinusoids in Gaussian noise}
\label{app:sinusoid_likelihood}
\jam{[to do! from Andrew: ``Write a short appendix proving that maximizing the likelihood exp(-n||n/2) with respect to A given a signal A cos (wt + phi) is the Fourier transform. Classic sig proc result useful for teachers to know.'']}

@@

\han{to do! pseudo-code here}
\jam{[Hannah has already explained a lot about Viterbi, at this point is it worth just citing existing documentation for further details? I used Wikipedia in writing code, should we cite that? Andrew mentions going through the practical aspects, e.g. log to avoid underflow, is that worth writing up? My thought is only that all of this is technically available in the open source.]}
%We consider a spectrogram grid where each element in the grid is the Fourier amplitude of a particular frequency at a particular time. These are normalised between $(0, 1)$ to use as multiplicative weights. In fact, we take the logarithm and so the sum of all weights to avoid underflow. The best path has the highest weight among all paths.

@@

Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram. 
It is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}. A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}. This design was based off \jam{standard examples (I have no idea where I or Alex originally got the designs from, I'm confident though that they are common enough to not need citing given that we also give the diagram)} of photo-detectors, connecting an ADC to a Pi, and Sallen-Key filters.

### 2020-06-30 paper/ifo-complexAudio.tex
winter comments part 2

% Optical microphone
% aim of optical microphone: play back injected sounds
A natural successor to simple tomes is complex audio, such as music and speech. 
Complex audio signals are not quasimonochromatic, unlike continuous gravitational wave signals, so the Viterbi algorithm as implemented in Section~\ref{sec:viterbi_wandering} is not applicable directly.
Instead, in this section, we explore how to use the interferometer as an `optical microphone' to detect and play back complex audio signals, using light to capture sound. 
In doing so, we experiment with a hierarchy of passive filters which suppress noise yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter tuned to sinusoidal signals in Section~\ref{sec:single_tone} and \han{to do: Appendix~\ref{}}. 
The device has precedence in the laser microphones~\cite{laser_microphone} used in the defence industry which operate on a variety of related principles.
It is also analogous to the detection of unmodelled gravitational-wave signals, e.g. broadband bursts from supernovae~\cite{}\han{to do - add citation}. 
%It is also analogous to the detection of unmodelled gravitational wave signals, e.g. broadband bursts from supernovae. [refs; my last sentence can be improved]

@@

The voltage signal from the photo-detector is captured by a MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2. 
Together, these sample the signal at roughly $16\,{\rm kHz}$.

@@

The optical microphone is tested on recordings of speech from different people and on music ranging from simple melodies to full track songs. 
Each recording is around one minute long with processing applied to ten second segments. 
For each observation, the audio is played through the speaker whilst the Pi script records the signal from the photodiode. 
During recordings, care is taken to ensure minimal activity around the interferometer to reduce environmental noise. 
The timeseries is then directly converted to a .wav file and played as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python} as before.

We find that initial recordings are extremely noisy with a loud bass hum throughout.
Each $10\,{\rm s}$ Fourier spectrum shows dominant AC mains noise with power ranging from the fundamental of the $50\,{\rm Hz}$ mains power up to and beyond its $8$th harmonic.
We find that the mains power signal is present in data recorded without an injected audio signal as seen in the power spectral density in Figure~\ref{fig:psd_noise}.
The mains power is also present in data taken with the photodiode in complete darkness with a similar but weaker distribution as that shown in Figure~\ref{fig:psd_noise}).
The origin of the mains power signal is likely to be from the photodiode signal, or environmental noise from the rooms lighting and air conditioning, or a combination.

@@

We note that simply zeroing the frequency channel of each $50\,{\rm Hz}$ in the frequency spectrum is unsuccessful. 
Effectively, this multiplies the spectrum by a rectangular comb filter. 
It does remove the mains harmonics, however it audibly ruins the rest of the signal. 
Applying a filter in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter. 
The inverse Fourier transform of a rectangular comb filter (a set of boxcars) is a complicated sinc-like function, which significantly corrupts the signal.

%A smoother comb filter significantly attenuates the spectrum between each harmonic so that the signal is no longer audible.

@@

\begin{figure}
	\includegraphics[width=0.49\textwidth]{figures/butterworth_150_3000-cropped.pdf}
	\caption{Butterworth bandpass filter frequency response, any amplitude beyond $-3\,{\rm dB}$ is significant attenuation (half power). The green vertical lines show the band limits of $150\,{\rm Hz}$ and $3\,{\rm kHz}$. Note the flat response within the band.}
	\label{fig:butterworth}
\end{figure}

Band-passing the signal similarly does not address the remaining harmonics above $100\,{\rm Hz}$. 
However, it does suppress other noise at higher frequencies. 
The best performing band-pass filter we find is a fifth order ($n = 5$) Butterworth filter over the range $150\,{\rm Hz}$--$3\,{\rm kHz}$ (beyond $2\,{\rm kHz}$ speech information content drops significantly~\cite{speech_intelligibility}). 
For a low-pass Butterworth filter with cut-off frequency $\omega_c$ and gain $\epsilon$, the frequency response (attenuation at each frequency) is,
\begin{equation}
\label{eq:butterworth}
H(\omega) = \frac{1}{\sqrt{1+\epsilon^2 (\omega/\omega_c)^{2n}}},
\end{equation}
as a function of frequency $\omega$. 
To construct a Butterworth band-pass filter, the low-pass filter in Eq.~\ref{eq:butterworth} is combined with a high-pass filter of similar form. 
The response of the filter used is shown in Figure~\ref{fig:butterworth}. 
The Butterworth filter performs the well as it is optimised to be `maximally flat' within the band region (prioritising condition (ii) above).
This filter improves on the above simpler filters but recordings of speech still have low intelligibility.
They can be recognised as speech without being understood.
Looking forward to Sec.~\ref{sec:logmmse}, results for the Butterworth filter are shown in combination with a speech enhancement technique.

@@

We also consider a combined notch filter and Wiener filter. 
Timeseries and frequency spectrum results are shown in Figs.~\ref{fig:timeNotchWiener} and ~\ref{fig:freqNotchWiener} respectively. 
The original signal is identical to that shown in Fig.~\ref{fig:noiseSubtract}.
The first and second panels of Fig.~\ref{fig:timeNotchWiener} show the original audio signal and the recording after passing thought the optical microphone (i.e. the timeseries of the data shown in the first two panels of Fig.~\ref{fig:noiseSubtract}).

A notch filter attenuates signals in specific frequency bands. 
We apply a $15$ cascaded notched filter to remove the mains noise and its harmonics up to $750\,{\rm Hz}$. 
\han{should we include here a mathematical description of the $15$ cascade notch filter?}
The timeseries and frequency spectrum results are shown in the third panel of Fig.~\ref{fig:timeNotchWiener} and the first panel of Fig.~\ref{fig:freqNotchWiener}. 
We seen that the mains harmonics are significantly attenuated in comparison to the analysis in Fig.~\ref{fig:noiseSubtract}.

The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 
It minimizes the mean square error $E(\hat{x}-x)^2$ between $\hat{x}$ and the actual signal $x$~\citep{verhoeven2009robust}.
\han{should we add more details here too?}
The timeseries and frequency spectrum results of the Wiener filter are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively. 
The timeseries is visibly cleaner, however there is a strong noise hum.

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{figures/filter_timeseries_aa_melatos-cropped.pdf}
	\caption{Time series of the optical microphone recording of an adult male voice (saying ``A cathode ...'') shown before (top panel) and after (bottom panel) the logMMSE estimator is applied. One second of the one minute recording is shown. The rise at the start of the bottom panel is expected when filtering a finite signal.}
	\label{fig:logMMSE_timeseries}
\end{figure*}

\begin{figure*}
	\includegraphics[width=\textwidth]{figures/filter_spectrum_aa_melatos-cropped.pdf}
	\caption{Spectrum of optical microphone recording of the same adult male voice in Fig.~\ref{fig:logMMSE_timeseries} shown before (top panel) and after (bottom panel) the logMMSE estimator is applied. The spectrum shows up to $2\,{\rm kHz}$ and to amplitude $2$ to see greater detail.}
	\label{fig:logMMSE_spectrum}
\end{figure*}

>>>>

% Optical microphone
% aim of optical microphone: play back injected sounds
A natural successor to simple tomes is complex audio, such as music and speech. 
Complex audio signals are not quasi-monochromatic, unlike continuous gravitational wave signals, so the Viterbi algorithm as implemented in Section~\ref{sec:viterbi_wandering} is not applicable directly.
Instead, in this section, we explore how to use the interferometer as an `optical microphone' to detect and play back complex audio signals, using light to capture sound. 
In doing so, we experiment with a hierarchy of passive filters which suppress noise yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter tuned to sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood} \han{to do}. 
The device has precedence in the laser microphones~\cite{laser_microphone} used in the defence industry which operate on a variety of related principles.
It is also analogous to the detection of unmodelled gravitational-wave signals, e.g. broadband bursts from supernovae~\cite{}\han{to do - add citation}. 
%It is also analogous to the detection of unmodelled gravitational wave signals, e.g. broadband bursts from supernovae. [refs; my last sentence can be improved]

@@

The voltage signal from the photo-detector is captured by a MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2. \jam{[Hannah is right, the Pi first appears here. If we want to comment on its convenience and usage this is the place.]}
Together, these sample the signal at roughly $16\,{\rm kHz}$.

@@

The optical microphone is tested on recordings of speech from different people and on music ranging from simple melodies to full track songs. 
Each recording is around one minute long with \jam{processing [this is poor, each recording was functionally ten seconds long, the fact that the sources were minutes long doesn't really matter]} applied to ten second segments. 
For each observation\jam{[again, poor wording, this means that for each 10 s recording we did the following]}, the audio is played through the speaker whilst the Pi script records the signal from the photodiode. 
During recordings, care is taken to ensure minimal activity around the interferometer to reduce environmental noise. 
The timeseries is then directly converted to a .wav file and played as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python} as before.

We find that initial recordings \jam{[Hannah's interpretation is correct here: "We find that the raw output of the optical microphone, collected without special precautions except anti-aliasing, is noisy, with a loud..."?]} are extremely noisy with a loud bass hum throughout.
Each $10\,{\rm s}$ Fourier spectrum shows dominant AC mains noise with power ranging from the fundamental of the $50\,{\rm Hz}$ mains power up to and beyond its $8$th harmonic. \jam{[non-linear appendix, Andrew says to cite Feynman lectures on non-linear oscillations]}
We find that the mains power signal is present in data recorded without an injected audio signal as seen in the power spectral density in Figure~\ref{fig:psd_noise}.
The mains power is also present in data taken with the photodiode in complete darkness with a similar but weaker distribution as that shown in Figure~\ref{fig:psd_noise}).
The origin of the mains power signal is likely to be from the photodiode signal, or environmental noise from the rooms lighting and air conditioning, or a combination. 
\jam{[No idea about the hump at 750 Hz. It is near to the main speech content, but is too broad for me to have any idea]}

@@

We note that simply zeroing the frequency channel of each $50\,{\rm Hz}$ in the frequency spectrum is unsuccessful. 
Effectively, this multiplies the spectrum by a rectangular comb filter. 
It does remove the mains harmonics, however it audibly ruins the rest of the signal. 
Applying a filter in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter. 
The inverse Fourier transform of a rectangular comb filter (a set of boxcars) is some combination of sinc functions, which significantly corrupt the signal.

%A smoother comb filter significantly attenuates the spectrum between each harmonic so that the signal is no longer audible. 
\jam{[re:comb filters. When I tried to comb filter it either was too sharp and so jumbled the whole signal in convolution. Or if smooth attenuated the signal away into noise. At least, that's what I remember. Changrong has a comb filter section anyway so I am happy dropping my unsuccessful attempts.]}

@@

\begin{figure}
	\includegraphics[width=0.49\textwidth]{figures/butterworth_150_3000-cropped.pdf}
	\caption{Butterworth bandpass filter frequency response, any amplitude beyond $-3\,{\rm dB}$ is significant attenuation (half power). The green vertical lines show the band limits of $150\,{\rm Hz}$ and $3\,{\rm kHz}$. Note the flat response within the band.
	\jam{[to do: change green lines to dashed and red. add panel with PSD after applying Butterworth bandpass]}}
	\label{fig:butterworth}
\end{figure}

Band-passing the signal similarly does not address the remaining harmonics above $100\,{\rm Hz}$. 
However, it does suppress other noise at higher frequencies. 
The best performing band-pass filter we find is a fifth order ($n = 5$) Butterworth filter over the range $150\,{\rm Hz}$--$3\,{\rm kHz}$ (beyond $2\,{\rm kHz}$ speech information content drops significantly~\cite{speech_intelligibility}). 
For a low-pass Butterworth filter with cut-off frequency $\omega_c$ and gain $\epsilon$, the frequency response (attenuation at each frequency) is,
\begin{equation}
\label{eq:butterworth}
H(\omega) = \frac{1}{\sqrt{1+\epsilon^2 (\omega/\omega_c)^{2n}}},
\end{equation}
as a function of frequency $\omega$. 
\jam{[Andrew wanted a quantitative comment about what the Butterworth does to the PSD, what do I use?]}
To construct a Butterworth band-pass filter, the low-pass filter in Eq.~\ref{eq:butterworth} is combined with a high-pass filter of similar form. 
The response of the filter used is shown in Figure~\ref{fig:butterworth}. 
The Butterworth filter performs the well as it is optimised to be `maximally flat' within the band region (prioritising condition (ii) above).
This filter improves on the above simpler filters but recordings of speech still have low intelligibility.
They can be recognised as speech without being understood.
Looking forward to Sec.~\ref{sec:logmmse}, results for the Butterworth filter are shown in combination with a speech enhancement technique.

@@

We also consider a combined notch filter and Wiener filter.
Timeseries and frequency spectrum results are shown in Figs.~\ref{fig:timeNotchWiener} and ~\ref{fig:freqNotchWiener} respectively. 
The original signal is identical to that shown in Fig.~\ref{fig:noiseSubtract}.
The first and second panels of Fig.~\ref{fig:timeNotchWiener} show the original audio signal and the recording after passing through the optical microphone (i.e. the timeseries of the data shown in the first two panels of Fig.~\ref{fig:noiseSubtract}).

A notch filter attenuates signals in specific frequency bands. 
We apply a $15$ cascaded notched filter to remove the mains noise and its harmonics up to $750\,{\rm Hz}$. 
\han{should we include here a mathematical description of the $15$ cascade notch filter?}
\jam{[Yes, I have no idea what a cascaded notched filter is and I'm an author!]}
The timeseries and frequency spectrum results are shown in the third panel of Fig.~\ref{fig:timeNotchWiener} and the first panel of Fig.~\ref{fig:freqNotchWiener}. 
We seen that the mains harmonics are significantly attenuated in comparison to the analysis in Fig.~\ref{fig:noiseSubtract}.

The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 
It minimizes the mean square error $\mathbb{E}[(\hat{x}-x)^2]$ between $\hat{x}$ and the actual signal $x$~\citep{verhoeven2009robust}.
\han{should we add more details here too?} \jam{[seconded]}
The timeseries and frequency spectrum results of the Wiener filter are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively. 
The timeseries is visibly cleaner, however there is a strong noise hum.

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{figures/filter_timeseries_aa_melatos-cropped.pdf}
	\caption{Time series of the optical microphone recording of an adult male voice (saying ``A cathode ...'') shown before (top panel) and after (bottom panel) the logMMSE estimator is applied. One second of the one minute recording is shown. The rise at the start of the bottom panel is expected when filtering a signal of finite duration.\jam{[add more panels showing original time series, like in Changrong's figures]}}
	\label{fig:logMMSE_timeseries}
\end{figure*}

\begin{figure*}
	\includegraphics[width=\textwidth]{figures/filter_spectrum_aa_melatos-cropped.pdf}
	\caption{Spectrum of optical microphone recording of the same adult male voice in Fig.~\ref{fig:logMMSE_timeseries} shown before (top panel) and after (bottom panel) the logMMSE estimator is applied. The spectrum shows only a detail of the total frequency domain (there is little activity at higher frequencies) and has truncated peaks at amplitude 2 which otherwise dominate the plot.}
	\label{fig:logMMSE_spectrum}
\end{figure*}

### 2020-06-30 paper/ifo-conclusions.tex
adding some future work text from Changrongs notes
winter comments part 2

In this paper we present an educational demonstration in which an \han{inexpensive}, table top Michelson interferometer serves as an analogue to a gravitational-wave detector. 
Vibrations produced by gravitational-wave signals are simulated with sound recordings played though a speaker fixed to the back of one of the interferometer mirrors. 
We consider quasimonochromatic signals as an analogue to continuous wave signals as well as short segments of speech and music which are (more distant) analogues of unmodelled gravitational-wave bursts, e.g. from supernovae. 
% This serves to demonstrate signal processing techniques.

@@

In Section~\ref{sec:complex_audio} we demonstrate recovery of more complex signals of music and speech. 
Substituting the webcam for the increased frequency response of a photodiode read by a Raspberry Pi, we transform the interferometer into an optical microphone.
We demonstrate that the optical microphone is capable of playing back simple recordings well but that it produces unintelligible recordings of speech. 
Mains hum at $50\,{\rm Hz}$ and its harmonics dominates the photodiode output. 
This offers students the opportunity to explore speech enhancement techniques to remove it, including the well performing logMMSE estimator. 
Further avenues of exploration include a more complete model of the optical microphone transfer function, which may lead to improvements in intelligibility.

>>>>

In this paper we present an educational demonstration in which an \han{inexpensive} \jam{[see comments in introduction]}, table top Michelson interferometer serves as an analogue to a gravitational-wave detector. 
Vibrations produced by gravitational-wave signals are simulated with sound recordings played though a speaker fixed to the back of one of the interferometer mirrors. 
We consider quasi-monochromatic signals as an analogue to continuous wave signals as well as short segments of speech and music which are (more distant) analogues of unmodelled gravitational-wave bursts, e.g. from supernovae. 
% This serves to demonstrate signal processing techniques.

@@

In Section~\ref{sec:optical_microphone} we demonstrate recovery of more complex signals of music and speech. 
Substituting the webcam for the increased frequency response of a photodiode read by a Raspberry Pi, we transform the interferometer into an optical microphone.
\jam{[Confirming Hannah's comment, the Pi was only necessary to read the photodiode, the webcam video I just saved onto PC and analysed directly]}
We demonstrate that the optical microphone is capable of playing back simple recordings well but that it produces unintelligible recordings of speech. 
Mains hum at $50\,{\rm Hz}$ and its harmonics dominates the photodiode output. 
This offers students the opportunity to explore speech enhancement techniques to remove it, including the well performing logMMSE estimator. 
Further avenues of exploration include a more complete model of the optical microphone transfer function, which may lead to improvements in intelligibility.

### 2020-06-30 paper/ifo-futureWork.tex
winter comments part 2

A second improvement is to measure the full model of the transfer function of the system. 
Theoretically, with the correct transfer function, signal recovery is limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
The transfer function starts from the voltage sent to the speaker and end with the value recorded by the Pi, and include any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation Appendix~\ref{app:intensity_derivation}, and the photodiode reading. 
A schematic of the full optical microphone transfer function is shown in Figure~\ref{fig:pipeline_highlighted}.

@@

Other traditional signal processing methods for extracting spectrogram like wavelet transform (WT) \citep{nason1995stationary} can also be implemented and compared with HMM. WT provides both time and frequency information, making it easier to pinpoint the origin of noise respect to time. 
In Refs. \cite{tufekci2000feature,agbinya1996discrete}, WT was also proposed for speech recognition.

@@

\begin{figure*}
	\includegraphics[width=.99\textwidth]{figures/pipeline_highlighted.pdf}
	\caption{
Flowchart for the optical microphone transfer function. 
The highlighted stage is analysised in Appendix XX, the other stages are left to future work. 
}
	\label{fig:pipeline_highlighted}
\end{figure*}

>>>>

A second improvement is to measure the full model of the transfer function of the system. 
With perfect knowledge of the transfer function, signal recovery is limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
\jam{[check with Rob et al. about the veracity of this]}
The transfer function starts from the voltage sent to the speaker and ends with the voltage recorded by the Pi. It includes any non-linearities in the speaker itself, the speaker-mirror coupling, the separation-intensity relation Appendix~\ref{app:intensity_derivation}, and the photodiode reading. 
A schematic of the full optical microphone transfer function is shown in Figure~\ref{fig:pipeline_highlighted}.

@@

Other traditional signal processing methods for extracting spectrogram like wavelet transform (WT) \citep{nason1995stationary} can also be implemented and compared with HMM. WT provides both time and frequency information, making it easier to pinpoint the origin of noise respect to time. 
In Refs.~\cite{tufekci2000feature,agbinya1996discrete}, WT was also proposed for speech recognition.

@@

\begin{figure*}
	\includegraphics[width=.99\textwidth]{figures/pipeline_highlighted.pdf}
	\caption{
Flowchart for the optical microphone transfer function. 
The highlighted stage is analysed in Appendix \jam{XX}, the other stages are left to future work. 
\jam{[to do: change grey to something more readable and reconsider labels. add editable version to git]}
}
	\label{fig:pipeline_highlighted}
\end{figure*}

### 2020-06-30 paper/ifo-introduction.tex
winter comments part 2

% sound analogy
In this paper, we present an \han{inexpensive} table-top experiment, where a standard Michelson interferometer plays the role of a `toy' gravitational-wave detector, which detects acoustic vibrations (such as music) instead of gravitational waves. 
As an undergraduate lab experiment this demonstration can be used to teach topics ranging from gravitational-wave detection and analysis to electronics, signal processing, and speech enhancement. 
It gives students in courses such as physics and electrical engineering the freedom to explore the response of an accessible yet nontrivial optomechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}.
The demonstration can also be adapted for use as an outreach tool to explain gravitational-wave science to a general audience.

>>>>

% sound analogy
In this paper, we present an \han{inexpensive} \jam{["Inexpensive" is relative. The experiment is cheap compared to LIGO, but out of budget for a high-school. Suggestions: "modest", "simple", or just cut the adjective.]} table-top experiment, where a standard Michelson interferometer plays the role of a `toy' gravitational-wave detector, which detects acoustic vibrations (such as music) instead of gravitational waves. 
As an undergraduate lab experiment this demonstration can be used to teach topics ranging from gravitational-wave detection and analysis to electronics, signal processing, and speech enhancement. 
It gives students in courses such as physics and electrical engineering the freedom to explore the response of an accessible yet nontrivial opto-mechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}.
The demonstration can also be adapted for use as an outreach tool to explain gravitational-wave science to a general audience.

### 2020-06-30 paper/ifo-singleTone.tex
winter comments part 2

%The resulting data is a video capture of the whole interference pattern. 
%As only the frequency of the interference ring motion is needed, w
We take data from a single off-centre pixel indicated by the red dot in Fig.~\ref{fig:interference_pattern}. 
The green-channel of the video is used as an approximation to the total intensity as the laser produces monochromatic green light (at $532\,{\rm nm}$).
As an initial inspection, we play the timeseries of the intensity using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
The two tones sound similar by ear and don't produce any noticeable beating. 
The result is quantified in Fig.~\ref{fig:webcam_spectrum}. 
A tone of $2.09\,{\rm Hz}$ is played through the speaker and the data recorded via webcam. 
Figure~\ref{fig:webcam_spectrum} shows the Fourier transform of the recording.
\han{[to do - rewrite once James has the new numbers (also Fig 3 caption)] 
We measure a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$ (where the error represents the $1\sigma$ uncertainty).
The demonstration is able to recover a single frequency tone to within two standard deviations. }
%We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
%The two tones played together sound similar and don’t produce any noticeable beating.
%To quantify the precision, we recorded a signal from an injected tone of $2.09\,{\rm Hz}$. 
%From the Fourier transform shown in Figure~\ref{fig:webcam_spectrum}, we measured a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$, a result within two standard deviations. 
%The experiment is able to recover a single tone confidently. 
%This showed that the experiment observes a single tone confidently.

\begin{figure}
	\includegraphics[width=.49\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
	\caption{\label{fig:webcam_spectrum}
Injected tone at a constant frequency: Fourier amplitude of the intensity time series recorded by the webcam. 
%Single injected tone results. 
%The Fourier amplitude of the intensity timeseres recorded by webcam. 
The frequency of the injection is $2.09\,{\rm Hz}$. 
The recovered frequency peak is at $(2.099\pm 0.008)\,{\rm Hz}$. 
\han{[to do: add new numbers from James]}
%The peak (with amplitude $17.5$) has a frequency was measured at $(2.099\pm 0.008) \mathrm{Hz}$. The actual maximum is achieved at $0\,{\rm Hz}$, but corresponds to the signal having a non-zero average value, so we don’t display or consider it.
}

>>>>

%The resulting data is a video capture of the whole interference pattern. 
%As only the frequency of the interference ring motion is needed, w
We take data from a single off-centre pixel indicated by the red dot in Fig.~\ref{fig:interference_pattern}. 
The green-channel of the video is used as an approximation to the total intensity as the laser produces monochromatic green light (at $532\,{\rm nm}$).
As an initial inspection, we play the timeseries of the intensity using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
The two tones sound similar by ear and don't produce any noticeable beating. 
The result is quantified in Fig.~\ref{fig:webcam_spectrum}. 
A tone of $2.09\,{\rm Hz}$ is played through the speaker and the data recorded via webcam. 
Figure~\ref{fig:webcam_spectrum} shows the Fourier transform of the recording.
\han{[to do - rewrite once James has the new numbers (also Fig 3 caption)] 
We measure a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$ (where the error represents the $1\sigma$ uncertainty).
The demonstration is able to recover a single frequency tone to within two standard deviations. }
\jam{[For the single tone result, the FWHM is 0.033 Hz. This puts the injected 2.09 Hz well within the uncertainty in the measured value of 2.099 Hz.]}
%We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
%The two tones played together sound similar and don’t produce any noticeable beating.
%To quantify the precision, we recorded a signal from an injected tone of $2.09\,{\rm Hz}$. 
%From the Fourier transform shown in Figure~\ref{fig:webcam_spectrum}, we measured a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$, a result within two standard deviations. 
%The experiment is able to recover a single tone confidently. 
%This showed that the experiment observes a single tone confidently.

\jam{[re: quantifying the non-linear response, the first two harmonics (at 4.19 and 6.28 Hz) are 0.103 and 0.098 the height of the main peak respectively.]}

\begin{figure}
	\includegraphics[width=.49\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
	\caption{\label{fig:webcam_spectrum}
Injected tone at a constant frequency: Fourier amplitude of the intensity time series recorded by the webcam. 
%Single injected tone results. 
%The Fourier amplitude of the intensity timeseres recorded by webcam. 
The frequency of the injection is $2.09\,{\rm Hz}$. 
The recovered frequency peak is at $(2.099\pm 0.008)\,{\rm Hz}$. 
\han{[to do: add new numbers from James]}
\jam{[For the single tone result, the FWHM is 0.033 Hz. This puts the injected 2.09 Hz well within the uncertainty in the measured value of 2.099 Hz.]}
%The peak (with amplitude $17.5$) has a frequency was measured at $(2.099\pm 0.008) \mathrm{Hz}$. The actual maximum is achieved at $0\,{\rm Hz}$, but corresponds to the signal having a non-zero average value, so we don’t display or consider it.
}

### 2020-06-30 paper/ifo-tableTopGWs.tex
winter comments part 2

A commonly-used analogy in explaining gravitational-wave science is sound. 
Signals of binary black hole and neutron star mergers can be converted into audio as an aid to explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}~\footnote{When introducing the acoustic analogy to lay audiences, it ins important to emphasize that gravitational waves are not sound. For example, gravitational waves can propagate in a vacuum, whereas sound cannot.}.
Sound vibrations provide a simple means to move the components of a demonstration interferometer. 
The vibrations move the mirrors, changing the length of the interferometer arms, and the resulting interference pattern. 
We use sound throughout this work as a means of simulating gravitational-wave-like signals and the data analysis and signal processing techniques used to detect them.

@@

>>>>

A commonly-used analogy in explaining gravitational-wave science is sound. 
Signals of binary black hole and neutron star mergers can be converted into audio as an aid to explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}~\footnote{When introducing the acoustic analogy to lay audiences, it is important to emphasize that gravitational waves are not sound. For example, gravitational waves can propagate in a vacuum, whereas sound cannot.}.
Sound vibrations provide a simple means to move the components of a demonstration interferometer. 
The vibrations move the mirrors, changing the length of the interferometer arms, and the resulting interference pattern. 
We use sound throughout this work as a means of simulating gravitational-wave-like signals and the data analysis and signal processing techniques used to detect them.

@@

\jam{[I don't know what these other designs are (haven't seen the references), is this something that the Adelaide people would have used in their design?]}
